---
title: "MasterControl QAL Performance — Exploratory Data Analysis"
author: "Group 4"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

# Introduction

The MasterControl QAL Performance dataset contains 16,644 qualified leads (QALs) from 2024–2025.

MasterControl sells two product families:

- Qx (Quality) — established, 19.7% progression  
- Mx (Manufacturing) — newer, 12.7% progression  

The goal of this EDA is to:

- Identify which **account types** convert best for Mx  
- Identify which **job titles** convert best for Mx  
- Compare Mx vs Qx performance across segments  
- Find cross-sell opportunities  
- Provide actionable recommendations  

# Description of the data

The dataset includes 14 fields covering:

- Lead outcomes  
- Account characteristics  
- Contact characteristics  
- Product information  
- Lead source  
- Time and identifiers  

Data quality is strong across most fields except titles (29% missing).

# Data import and initial checks

```{r data import}
library(tidyverse)
library(janitor)
library(lubridate)
library(skimr)
library(naniar)
library(ggplot2)

qal <- read_csv("QAL Performance for MSBA.csv") %>% 
  clean_names()

str(qal)
```

# Missing data

Let's start first by analyzing the missing values from the data.

```{r}
skim(qal)
```

The summary shows that roughly 40% of values are missing in:
- acct_manufacturing_model (business model)
- acct_primary_site_function (what the site manufactures)
- contact_lead_title (lead title)
To visualize the pattern of missingness, we generate a heatmap and compute missing counts per column:

```{r heat map}
vis_miss(qal)

qal_missing <- qal |>
  summarise(across(everything(), ~sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "missing") |>
  arrange(desc(missing))

qal_missing

qal |> 
  summarise(
    missing_titles = mean(is.na(contact_lead_title))
  )
```

Key Findings
- Contact lead titles: ~38% missing → treat “Missing” as its own category
- next_stage__c: 483 missing → these rows are excluded when constructing the target variable
- All other fields: <1% missing → no special handling required

# Target variable construction

```{r target variable}
qal <- qal |>
  mutate(
    progressed = case_when(
      next_stage_c %in% c("SQL", "SQO", "Won") ~ 1,
      next_stage_c %in% c("Recycled", "Exclude", "Lost", "MQL", "INQ") ~ 0,
      TRUE ~ NA_real_
    )
  ) |>
  filter(!is.na(progressed))

qal |>
  count(progressed) |>
  mutate(pct = n / sum(n))
```

# Exploratory visualizations and summary tables

## Overall lead outcomes

```{r}
qal |>
  count(progressed) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(x = factor(progressed), y = n, fill = factor(progressed))) +
  geom_col() +
  theme_minimal()
```

## Product comparison (Qx vs Mx)

```{r}
product_summary <- qal |>
  group_by(solution_rollup) |>
  summarise(
    leads = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

product_summary
```

## Account-level EDA

### Industry

```{r}
industry_summary <- qal |>
  group_by(acct_target_industry, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

industry_summary
```

### Manufacturing model

```{r}
manufacturing_summary <- qal |>
  group_by(acct_manufacturing_model, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  ) |>
  arrange(desc(n))

manufacturing_summary
```

### Primary site function

```{r}
site_function_summary <- qal |>
  group_by(acct_primary_site_function, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  ) |>
  arrange(desc(n))

site_function_summary
```

### Company size

```{r}
size_summary <- qal |>
  group_by(acct_tier_rollup, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

size_summary
```

### Territory

```{r}
territory_summary <- qal |>
  group_by(acct_territory_rollup, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

territory_summary
```

## Title-level EDA

### Extract seniority and function keywords

```{r}
qal <- qal |>
  mutate(
    title_lower = tolower(contact_lead_title),
    has_quality = str_detect(title_lower, "quality"),
    has_ops = str_detect(title_lower, "operat|manufactur"),
    has_reg = str_detect(title_lower, "regulat"),
    seniority = case_when(
      str_detect(title_lower, "chief|ceo|cfo|coo|cio|cto") ~ "C-Level",
      str_detect(title_lower, "vp|vice") ~ "VP",
      str_detect(title_lower, "director") ~ "Director",
      str_detect(title_lower, "manager") ~ "Manager",
      is.na(title_lower) ~ "Missing",
      TRUE ~ "Other"
    )
  )
```

### Conversion by seniority

```{r}
seniority_summary <- qal |>
  group_by(seniority, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

seniority_summary
```

### Conversion by functional flags

```{r}
function_summary <- qal |>
  group_by(has_quality, has_ops, has_reg, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

function_summary
```

## Combined account + title patterns

```{r}
combined_summary <- qal |>
  group_by(acct_manufacturing_model, seniority, has_ops, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  ) |>
  arrange(desc(progression_rate))

combined_summary
```

## Time-based patterns

```{r}
time_summary <- qal |>
  mutate(
    year = year(qal_cohort_date),
    month = floor_date(qal_cohort_date, "month")
  ) |>
  group_by(year, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

time_summary
```

## Lead source analysis

```{r}
priority_summary <- qal |>
  group_by(priority, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

priority_summary

channel_summary <- qal |>
  group_by(last_tactic_campaign_channel, solution_rollup) |>
  summarise(
    n = n(),
    progression_rate = mean(progressed),
    .groups = "drop"
  )

channel_summary
```

# Results section

Summaries we can include:

- Best account types for Mx  
- Best titles for Mx  
- Cross-sell opportunities  
- Expected impact  
- Actionable recommendations  

# Group member contribution

Describe each member’s contributions.