

```{r}
# Core libraries
library(tidyverse)
library(janitor)
library(lubridate)

# Modeling libraries
library(tidymodels)

# Set seed for reproducibility
set.seed(123)
```

```{r}
# Import data (update filename as needed)
qal_raw <- read_csv("Mastercontrol QAL Performance.csv")

qal <- qal_raw |>
  clean_names()

glimpse(qal)
```

# Target construction

```{r}
qal <- qal |>
  mutate(
    progressed = case_when(
      next_stage_c %in% c("SQL", "SQO", "Won") ~ 1,
      next_stage_c %in% c("Recycled", "Lost", "MQL", "INQ") ~ 0,
      TRUE ~ NA_real_
    )
  )

qal_model <- qal |>
  filter(!next_stage_c %in% c("Exclude")) |>
  filter(!is.na(progressed)) |>
  mutate(
    progressed = factor(progressed, levels = c(0, 1), labels = c("no", "yes"))
  )

qal_model |>
  count(progressed) |>
  mutate(pct = n / sum(n))
```

#Title feature engineering

```{r}
qal_model <- qal_model |>
  mutate(
    title_lower = tolower(contact_lead_title),
    seniority = case_when(
      str_detect(title_lower, "chief|ceo|cfo|coo|cio|cto") ~ "C-Level",
      str_detect(title_lower, "vp|vice") ~ "VP",
      str_detect(title_lower, "director") ~ "Director",
      str_detect(title_lower, "manager") ~ "Manager",
      is.na(title_lower) ~ "Missing",
      TRUE ~ "Other"
    ),
    has_quality = str_detect(title_lower, "quality"),
    has_ops     = str_detect(title_lower, "operat|manufactur|engineer"),
    has_reg     = str_detect(title_lower, "regulat")
  )
```

# Train/test split + resampling

```{r}
# Stratified split on progressed
set.seed(123)
qal_split <- initial_split(qal_model, prop = 0.8, strata = progressed)

qal_train <- training(qal_split)
qal_test  <- testing(qal_split)

# 5- or 10-fold CV, stratified
set.seed(123)
qal_folds <- vfold_cv(qal_train, v = 5, strata = progressed)
```

# 3.5 Recipe (missingness as informative, dummy encoding)
Adjust predictors as needed—this is a good starting point.

```{r}
qal_recipe <- recipe(progressed ~ acct_primary_site_function +
                       acct_manufacturing_model +
                       acct_target_industry +
                       acct_territory_rollup +
                       acct_tier_rollup +
                       solution_rollup +
                       seniority +
                       has_quality + has_ops + has_reg +
                       last_tactic_campaign_channel,
                     data = qal_train) |>

  # Convert logicals to factors BEFORE handling missingness
  step_mutate(
    has_quality = factor(has_quality, levels = c(FALSE, TRUE)),
    has_ops     = factor(has_ops, levels = c(FALSE, TRUE)),
    has_reg     = factor(has_reg, levels = c(FALSE, TRUE))
  ) |>

  # FIRST: catch unseen levels during CV
  step_novel(all_nominal_predictors()) |>

  # SECOND: convert NA to "unknown"
  step_unknown(all_nominal_predictors()) |>

  # One-hot encode AFTER novel/unknown handling
  step_dummy(all_nominal_predictors()) |>

  # Remove zero-variance predictors
  step_zv(all_predictors()) |>

  # Impute any remaining NA values
  step_impute_median(all_numeric_predictors()) |>
  step_impute_mode(all_nominal_predictors())
```

# 4 Models: specs + tuning patterns

## 4.1 Logistic regression (baseline, no penalty)

```{r}
log_reg_spec <- logistic_reg(mode = "classification") |>
  set_engine("glm")

log_reg_wf <- workflow() |>
  add_model(log_reg_spec) |>
  add_recipe(qal_recipe)

set.seed(123)
log_reg_res <- fit_resamples(
  log_reg_wf,
  resamples = qal_folds,
  metrics = metric_set(roc_auc, pr_auc, accuracy, sens, spec),
  control = control_resamples(save_pred = TRUE)
)

collect_metrics(log_reg_res)
```

The model shows strong sensitivity and PR AUC but low specificity, indicating it captures positives well while misclassifying many negatives.

#4.2 LASSO

```{r}
lasso_spec <- logistic_reg(
  mode = "classification",
  penalty = tune(),
  mixture = 1  # 1 = pure LASSO; you can also tune mixture for elastic net
) |>
  set_engine("glmnet")

lasso_wf <- workflow() |>
  add_model(lasso_spec) |>
  add_recipe(qal_recipe)

lasso_grid <- grid_regular(
  penalty(range = c(-4, 0)), # 10^-4 to 1
  levels = 20
)

set.seed(123)
lasso_res <- tune_grid(
  lasso_wf,
  resamples = qal_folds,
  grid = lasso_grid,
  metrics = metric_set(roc_auc, pr_auc, accuracy)
)

lasso_best <- select_best(lasso_res, metric = "roc_auc")
lasso_best

lasso_final_wf <- finalize_workflow(lasso_wf, lasso_best)
```
This is a small penalty meaning light regularization. In other words, many predictors still contribute and only the weakest coefficients are shrunk to zero. 

#4.3 Decision Tree

```{r}
tree_spec <- decision_tree(
  mode = "classification",
  cost_complexity = tune(),
  tree_depth      = tune(),
  min_n           = tune()
) |>
  set_engine("rpart")

tree_wf <- workflow() |>
  add_model(tree_spec) |>
  add_recipe(qal_recipe)

tree_grid <- grid_regular(
  cost_complexity(),
  tree_depth(),
  min_n(),
  levels = 4
)

set.seed(123)
tree_res <- tune_grid(
  tree_wf,
  resamples = qal_folds,
  grid = tree_grid,
  metrics = metric_set(roc_auc, pr_auc, accuracy)
)

tree_best <- select_best(tree_res, metric = "roc_auc")
tree_best

tree_final_wf <- finalize_workflow(tree_wf, tree_best)
```
This tuning selects an almost unpruned tree with deep splits, but the larger min_n keeps it from overfitting on tiny nodes.

#4.4 Random Forest

```{r}
rf_spec <- rand_forest(
  mode = "classification",
  mtry  = tune(),
  trees = 100,
  min_n = tune()
) |>
  set_engine("ranger", importance = "impurity")

rf_wf <- workflow() |>
  add_model(rf_spec) |>
  add_recipe(qal_recipe)

rf_grid <- grid_regular(
  mtry(range = c(5, 25)),
  min_n(),
  levels = 5
)

set.seed(123)
rf_res <- tune_grid(
  rf_wf,
  resamples = qal_folds,
  grid = rf_grid,
  metrics = metric_set(roc_auc, pr_auc, accuracy)
)

rf_best <- select_best(rf_res, metric = "roc_auc")
rf_best

rf_final_wf <- finalize_workflow(rf_wf, rf_best)
```

Each tree considers a reasonable number of predictors but avoids very small terminal nodes, keeping the forest stable.

# 4.5 Gradient Boosting

```{r}
xgb_spec <- boost_tree(
  mode = "classification",
  trees = 100,
  learn_rate = tune(),
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  mtry = tune()
) |>
  set_engine("xgboost")

xgb_wf <- workflow() |>
  add_model(xgb_spec) |>
  add_recipe(qal_recipe)

xgb_grid <- grid_latin_hypercube(
  learn_rate(),
  tree_depth(),
  min_n(),
  loss_reduction(),
  sample_size = sample_prop(),
  mtry(range = c(5, 25)),
  size = 10
)

set.seed(123)
xgb_res <- tune_grid(
  xgb_wf,
  resamples = qal_folds,
  grid = xgb_grid,
  metrics = metric_set(roc_auc, pr_auc, accuracy)
)

xgb_best <- select_best(xgb_res, metric = "roc_auc")
xgb_best

xgb_final_wf <- finalize_workflow(xgb_wf, xgb_best)
```

# 5 Model comparison

```{r}
log_reg_metrics  <- collect_metrics(log_reg_res)  |> mutate(model = "log_reg")
lasso_metrics    <- collect_metrics(lasso_res)    |> mutate(model = "lasso")
tree_metrics     <- collect_metrics(tree_res)     |> mutate(model = "tree")
rf_metrics       <- collect_metrics(rf_res)       |> mutate(model = "rf")
xgb_metrics      <- collect_metrics(xgb_res)      |> mutate(model = "xgb")

all_metrics <- bind_rows(
  log_reg_metrics,
  lasso_metrics,
  tree_metrics,
  rf_metrics,
  xgb_metrics
)

all_metrics |>
  filter(.metric %in% c("roc_auc", "pr_auc", "accuracy")) |>
  arrange(.metric, desc(mean))
```

Across all tuned models, LASSO emerges as the most reliable performer, consistently producing the highest accuracy regardless of penalty value. This suggests that a sparse linear model captures the underlying structure of the data better than more complex tree‑based or ensemble methods.

# Final fit /  test performance

```{r}
final_fit <- last_fit(
  lasso_final_wf,
  split = qal_split,
  metrics = metric_set(roc_auc, pr_auc, accuracy, sens, spec)
)

collect_metrics(final_fit)

final_preds <- collect_predictions(final_fit)
```

The final LASSO model performs about the same on the test set as it did during cross‑validation. It stays very strong on sensitivity and overall accuracy, with the same low‑specificity trade‑off, so its behavior is consistent and not overfit.